---
title: "Option pricing using Monte Carlo simulation"
output: html_notebook
---

The value of call option (ie. payoff) = max(Pt-K,0).

A common approach to pricing a derivative security is to employ Monte Carlo simulation. This involves simulating the prices of the underlying asset over a discrete grid of time points.

---

Code to simulate a single ray of asset prices at N equally-spaced time points under the assumption that the underlying security follows a Geometric Brownian motion. 

```{r}
# Simulation from a RW model (using a loop)
#this is just a single ray
mu <-0.01
sigma <- 0.03
N <- 100 #100 periods into the future
r <- rep(0,T)
P <- rep(0,T)
P0 <- 2

r[1] <- rnorm(1,mu,sigma)
P[1] <- P0*exp(r[1])
for (t in 2:N)
{
 r[t] <- rnorm(1,mu,sigma)
 P[t] <- P[t-1]*exp(r[t])
}
```


Code to obtain the stock prices of 3M Co. Ltd. for illustration purposes.

```{r}
library(tseries)
start="1990-01-01"
p3M=get.hist.quote("MMM", start=start, quote = "AdjClose",retclass="zoo",compression="m")

r3M = diff(log(p3M)) # continuously compounding returns 

mu = mean(r3M)
variance = var(r3M)

```

Code to simulate 10,000 rays of monthly price values one year into the future for stock MMM. The code also plots a histogram of the price of the stock one year into the future, the mean price as well as the variance. 

```{r}
mu <-mu
sigma <- sqrt(variance)
N <- 12 #12 months into the future
M <- 10000 # 10,000 rays 
P0 <- p3M[length(p3M)]
PT <- rep(0,M) #vector to store the predicted price 

for (m in 1:M){
  r <- rep(0,N)
  P <- rep(0,N)
  r[1] <- rnorm(1,mu,sigma)
  P[1] <- P0*exp(r[1])
  for (t in 2:N){
    r[t] <- rnorm(1,mu,sigma)
    P[t] <- P[t-1]*exp(r[t])
  }
  PT[m] <- P[N]
}

hist(PT)
mean(PT)
var(PT)
```

The code below is a Monte Carlo simulation to evaluate the distribution of the payoff 12 months into the future of a call option with strike price K = P0. 

```{r}
mu <-mu
sigma <- sqrt(variance)
N <- 12 #12 months into the future
M <- 10000 # 10,000 rays 
P0 <- p3M[length(p3M)]
PT <- rep(0,M) #vector to store the predicted price 
payoff <- rep(0,M) #vector to store call option payoff

for (m in 1:M){
  r <- rep(0,N)
  P <- rep(0,N)
  r[1] <- rnorm(1,mu,sigma)
  P[1] <- P0*exp(r[1])
  for (t in 2:N){
    r[t] <- rnorm(1,mu,sigma)
    P[t] <- P[t-1]*exp(r[t])
  }
  PT[m] <- P[N]
  payoff[m] <- max(0,(PT[m]-P0))
}

hist(payoff)
mean(payoff)

```

