 ---
title: "Pricing an American Put"
output: R Markdown
 ---
An American option can be exercised early (ie. before maturity). It is not profitable (in a risk neutral sense) to exercise an American call on a non-dividend paying security early. However, it may be profitable to exercise an American Put early. Hence, the Black-Scholes pricing formula will undervalue the American Put option and is not applicable in this scenario. The code below computes the price of an American Put option using a binomial tree. 
 ---
 
Function to calculate the price of the underlying asset

```{r}
Stock_Price <- function(P0,sig,Dt,N){
  X <- matrix(0,N+1,N+1)
  u = exp(sig*sqrt(Dt))
  d = exp(-sig*sqrt(Dt))
  for(i in 1:nrow(X)){
    for(j in 1:i){
      X[i,j] <- P0*(d^(i-j))*(u^(j-1))
    }
  }
  return(X)
}
```

Function to compute the risk neutral probability of the stock going up by a factor of u 

```{r}
risk_neutral_prob <- function(r,sig,Dt){
  u = exp(sig*sqrt(Dt))
  d = exp(-sig*sqrt(Dt))
  q = (exp(r*Dt)-d)/(u-d)
  return(q)
}
```

Function to compute the value of an American put (with strike price K) at each node of the tree 

```{r}
American_Put <- function(stock_price_tree,q,Dt,K,r){
  N <- nrow(stock_price_tree)-1
  F <- matrix(0,N+1,N+1)
  for (j in 1:(N+1)){
    F[N+1,j] <- max(0,K-stock_price_tree[N+1,j])
  }
  for(i in (nrow(F)-1):1){
    for(j in i:1){
      F[i,j] <- max(K-stock_price_tree[i,j], exp(-r*Dt)*(q*F[i+1,j+1]+(1-q)*F[i+1,j]))
    }
  }
  return(F)  
}
```

Function that gathers all the inputs from the above functions and returns the stock price and option value trees, the risk neutral probability q, the value of the time step size and the price of the American put option

```{r}
Binomial_Model <- function(P0,sig,T,N,K,r){
  Dt = T/N
  stock_price_tree <- Stock_Price(P0,sig,Dt,N)
  q <- risk_neutral_prob(r,sig,Dt)
  Option_value_tree <- American_Put(stock_price_tree,q,Dt,K,r)
  Option_price <- Option_value_tree[1,1]
  Output = NULL
  Output$Stock_price_tree = Stock_price_tree
  Output$Option_value_tree = Option_value_tree
  Output$risk_neutral_prob = q
  Output$time_step_size = Dt
  Output$Option_price = Option_price
  return(Output)
}
```

An example of pricing an American put with:
- intial price (P0) = $35
- strike price (K) = $36.50
- sigma (sig) = 0.28
- Time in years (T) = 9/12 (ie. 9 months)
- risk free rate of return (r) = 0.015

```{r}
Binomial_Model(35,0.28,9/12,3,36.5,0.015)
```

Plot option price against number of discrete periods

```{r}
price <- rep(0,47)
for (N in 3:49){
  temp <- Binomial_Model(35,0.28,9/12,N,36.5,0.015)
  price[N-2] <- temp[length(temp)]
}
plot((3:49),price, xlab = "Number of discrete periods", ylab = "Option Price", main = "Option price vs number of discrete periods")
```

